---
title: "M4L3_Homework_Assignment_Answers"
author: "Victor Gomes"
date: "March 22, 2016"
output: word_document
---

This homework assignment focuses on Expectation-maximization (EM).

## Additional packages needed
 
To run the code in M4L3_Homework_Assignment_Answers.Rmd you may need additional packages.

* If necessary install following packages.

`install.packages("fitdistrplus");`
`install.packages("mclust");`

```{r}
require(fitdistrplus)
require(mclust)
```


```{r}
data_url <- 'https://archive.ics.uci.edu/ml/machine-learning-databases/iris/iris.data'
iris_data <-read.table(url(data_url), sep=",")
colnames(iris_data)<-c("SepalLength","SepalWidth","PetalLength","PetalWidth","Class")
iris <- iris_data[1:4]
```

Yatish Jain's answer

Lets take a look at how the data is distributed using the Cullen and Frey Graph.
```{r}
descdist(iris$SepalLength, discrete = FALSE)
descdist(iris$SepalWidth, discrete = FALSE)
descdist(iris$PetalLength, discrete = FALSE)
descdist(iris$PetalWidth, discrete = FALSE)
```
From the above graph we can see that  Sepal length can be a beta distribution and Sepal width can be either lognormal or gamma distribution whereas Petal Length and Petal width are mostly uniform distrubutions.

###EM Based Clustering

* How did you choose a model for EM? Evaluate the model performance.

As the data seems to be a mixture of distributions we will be using EM model where we use the E step to sum the expectations of statistics and M step to maximise the linear function. As all the data doesn't belong to linear function it is good to use EM model here.

```{r}
fit<-Mclust(iris)
fit
summary(fit)
```

```{r}
# BIC
plot(fit,what="BIC")
plot(fit, what = "classification")
plot(fit, what = "uncertainty")
plot(fit, what = "density")
BIC = mclustBIC(iris)
summary(BIC)
```

* Cluster some of your data using EM based clustering that you also used for k-means, PAM, and hierarchical clustering. How do the clustering approaches compare on the same data?

EM clustering approach performs very similar to k-means, pam of heirarchical clustering.

