---
title: "M4L4_Homework_Assignment_Answers"
author: "Victor Gomes"
date: "March 22, 2016"
output: word_document
---

This homework assignment focuses on Association Rules.

## Additional packages needed
 
To run the code in M4L4_Homework_Assignment_Answers.Rmd you may need additional packages.

* If necessary install following packages.

`install.packages("arules");`
`install.packages("arulesViz");`

```{r}
require("arules")
require("arulesViz")
```

Retail Data Set:

```{r}
data_url <- 'http://fimi.ua.ac.be/data/retail.dat'
retail <- read.transactions(url(data_url))
summary(retail)
# look at the first five transactions
inspect(retail[1:5])
# Look at the frequency 
itemFrequency(retail[1:100, 1:10])
# plot the frequency 
itemFrequencyPlot(retail,support=0.1)
itemFrequencyPlot(retail,topN=20)
```

* Generate a set of 50 or so (non-redundant) rules.

```{r}
retail.rules<-apriori(retail)
summary(retail.rules)
```

Improving model performance.

```{r}
retail.rules.2 <- apriori(retail, parameter = list(support = 0.01, confidence = 0.5, minlen = 1))
retail.rules.2
# converting the rule set to a data frame
Rulesdataframe<- as(retail.rules.2, "data.frame")
str(Rulesdataframe)
```

Pruning redundant rules.

```{r}
subset.matrix <- is.subset(retail.rules.2, retail.rules.2)
subset.matrix[lower.tri(subset.matrix, diag=T)] <- NA
redundant <- colSums(subset.matrix, na.rm=T) >= 1
which(redundant)
# remove redundant rules
rules.pruned <- retail.rules.2[!redundant]
inspect(rules.pruned)
```
 

* Which rules make sense to you? Highlight the five best and five worst of your rule set. 

{} => {39}
{41}   => {48}
{38}   => {48}
{110}  => {38}
{170}  => {38}

The above rules look like the best since they have high support/confidence. 

{270}  => {48}
{2238} => {48
{1327} => {48}
{38}   => {48}
{170}  => {48}

The above rules look like the worst because of their low support and confidence. 

* How did you choose the level of support and confidence? 

Higher level of confidence and lower level of support is  preferable. So, we set support at 0.01 and confidence at 0.5.

* What is the lift and conviction of your best and worst rules? 

Lift of best rules -

{} => {39}  - 1.000000

{41}   => {48} - 1.262562

{38}   => {48} - 1.065772

{110}  => {38} - 5.513258

{170}  => {38} - 5.528821

Lift of worst rules -

{270}  => {48} - 1.154785

{2238} => {48 - 1.165139

{1327} => {48} - 1.134050

{38}   => {48} - 1.065772

{170}  => {48} - 1.051249

Conviction:

```{r}
conviction <- interestMeasure(rules.pruned, "conviction", transactions=retail)
rules.conviction<-as(rules.pruned, "data.frame")
rules.conviction<-data.frame(rules.conviction, conviction)
rules.conviction
```

Best rules -

{} => {39}  - 1.000000

{41}   => {48} - 1.316413

{38}   => {48} - 1.064068

{110}  => {38} - 33.329520

{170}  => {38} - 37.511499

Worst Rules -

{270}  => {48} - 1.165090

{2238} => {48 - 1.178099

{1327} => {48} - 1.139881

{38}   => {48} - 1.064068

{170}  => {48} - 1.049225


* Visualize your 50 association rules. Where do the best and worst end up in your plot? 

```{r}
plot(retail.rules.2)
plot(retail.rules.2, method="graph", control=list(type="items"))
plot(retail.rules.2, method="paracoord", control=list(reorder=TRUE))
```

* Does the model make sense?

The model makes sense. Each number represents one item in a supermarket and the rules we are valuable since the supermarket would benefit from using them in selling these items.


