---
title: "Module_01_Lesson_03 Solution"
author: "Romil Shah"
date: "05-15-2016"
---

```{r}
twitterdata<-read.csv(file="C:/Users/rams1/Desktop/DSCS6030/Module_01/M01_quasi_twitter.csv")
require(ggplot2)
```


##How is the data distributed? What are the summaries?

####(1) As per creation month-day-year
```{r echo=FALSE}
summary(twitterdata$created_at_month)
qplot(created_at_month, data=twitterdata, geom = "bar") + scale_x_discrete()
qplot(created_at_day,data=twitterdata,geom="histogram")
qplot(created_at_year,data=twitterdata,geom="histogram")
```

####(2) As per Twitter data by country
```{r}
summary(twitterdata$country)
```

####(3) As per Friends_count
```{r}
qplot(data=twitterdata, x=friends_count)
summary(twitterdata$friends_count)

```

####(4) As per gender
```{r}
qplot(data=twitterdata, x=gender)
summary(twitterdata$gender)

```

In the same way other columns can be used to classify the type of distribution of the data.


##Anomolies or Outliers

```{r}
boxplot(twitterdata$height)
boxplot(twitterdata$statuses_count)
boxplot(twitterdata$age)
```
There can be seen some outliers for all the box plots. 

```{r}
qplot(1,mobile_friends_count, data=twitterdata,xlab = "Outlier in friends_count")+geom_violin()
```
The violin plot also shows the outliers in the friends count.

##Can you identify:

1. Useful raw data transform
```{r}
qplot(x=log(followers_count+1,2), data=twitterdata)
```
This shows that the log followers are more between 5 to 12 in the log scale and there are very few beyond that range.

2. Data quality problems
* Age showing '0'
* Negative number in friends count

3. Oultiers
This was shown before that outliers lie in the places where there is data quality problem.

## Functional Relationships
``` {r}
ggplot(twitterdata, aes(x = age, y = friends_count, colour=gender)) + geom_point()
```
Males tend to have more friends count than females in each age group.

``` {r}
ggplot(twitterdata, aes(x = wage, y = education, colour=gender)) + geom_point()
```
Males tend to have more wage for their education compared to female counterparts.